Choosing harmonies based on average maximum pairwise distance of distributions

### 2020-05-07 algo_rank_harmony.Rmd
update algorithm

---
title: "Choosing harmonies based on average maximum pairwise distance of distributions"
author: "Sayani Gupta"
output:
  html_document:
    theme: united
    highlight: tango
    toc: true
---

@@

  a) Normal: Pekalska and Duin [23] assumed that Lp-distances between
feature vectors are normally distributed, if the feature values are independent, identically distributed, and have limited variance.
We can not assume that feature values are independent.

  b) Gamma: Ferencz et al. [7] have considered the Gamma distribution to model the L2-distances.
This is intuitive since if the reponse variable follows Normal distribution, then $L_2$ norm distance measure which is similar to taking squares of differences of Normal variables would follow a Chi-Squared distribution, a special case of Gamma distribution.

@@

# Algorithm

 1. Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_I$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_J$\} and $v$ is the measured variable. Hence for each combination ($A_i$, $B_j$), we have s subset of the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, I}$ and $\forall j = {1, 2, \dots, J}$

 2. There will be $J$ groups corresponding to each level $A_i$ of $C_1$. We compute the pairwise differences in distributions of these $J$ groups. Thus there will be  $k = \binom{J}{2}$ pairwise distances corresponding to each $A_i$. Let $D_1$,...,$D_k$ be these distances and let their maximum be $M_k$.

## Version 1 (Distances follow Gamma)

 3.1. Suppose the reponse variable follows Normal distribution. Then any distance measure which is similar to taking squares of differences of Normal variables would follow a Chi-Squared distribution, a special case of Gamma distribution. Hence if we proceed with the assumption that the $D_i$ distribution has Gamma distribution, then under EVT the maximum converges to Gumbel distribution and lim $P(M_k - b_k)/a_k)$ converges to a limiting Gumbel distribution where $b_k = (\frac{1}{\beta} (ln n + (\alpha - 1)ln ln n - ln \gamma(\alpha))$ and $a_k = \frac{1}{\beta}$. Here $\alpha$ and $\beta$ could be estimated using $(\mu/s)^2$ and $(s^2/\mu)$ respectively.

 4.1. For each facet, we have a normalised variable using Step 3.1. We can take the maximum across all facets as the final mesasure. But this maximum is not normalized against the number of facet levels. Or we can take the mean (to be discussed).

```{r}
read_rds("data/smart_harmony_gamma.rds")
```

@@

```{r}
read_rds("data/smart_harmony_ak.rds")
```

```{r}
smart_meter10 %>%
  filter(customer_id %in% c(10017936)) %>%
  prob_plot("day_week",
    "day_month",
    response = "general_supply_kwh",
    plot_type = "quantile",
    symmetric = TRUE,
    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
  ) +
  scale_y_sqrt() +
  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
  theme_minimal()
```

## Version 3 (Distances follow Normal)

 3.3. Let $D_{k} \sim N(\mu, \sigma^2)$. Then for $a_k = \mu − \sigma\Phi^{-1}(1/k)$ and $b_k = -\sigma/\Phi^{-1}(1/k)$, we have we have $P(\frac{M_k - a_k}{b_k} \leq z) \xrightarrow{} G(z)$, where $G$ is the standard Gumbel distribution. So normalization could be done  using $\frac{M_k - a_k}{b_k}$, where $\Phi^{-1}(1/k)$ could be approximated by $(1 - \frac{1}{k})$ sample quantile and $\mu$ and $\sigma$ by sample mean and standard deviation.

 4.3. We have been able to normalize for the number of levels of x-axis variable using Step 3.2. Now 
$\forall A_i, \quad i \in \{1, 2, \dots, I\}$, we have a normalised variable each of which follows a standard Gumbel distribution. Now we need to combine these normalised variables across different facets in a way so that the number of facet levels do not play a role. For this we can take the mean, median or any quantiles of these normalised variables. But the median (and any quantiles) of these Gumbel distributions grow at the rate of (some power of) $log n$ with $n \xrightarrow{} \infty$. Hence, median($N_i$)/log($I$) could serve as the final measure.

```{r}
library(readr)
read_rds("data/smart_harmony_musigma.rds")
```
```{r}
smart_meter10 %>%
  filter(customer_id %in% c(10017936)) %>%
  prob_plot("hour_day",
    "wknd_wday",
    response = "general_supply_kwh",
    plot_type = "quantile",
    symmetric = TRUE,
    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
  ) +
  scale_y_sqrt() +
  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
  theme_minimal()
```

@@

```{r}
library(readr)
read_rds("data/smart_harmony_n.rds")
```

```{r}
smart_meter10 %>%
  filter(customer_id %in% c(10017936)) %>%
  prob_plot("week_month",
    "wknd_wday",
    response = "general_supply_kwh",
    plot_type = "quantile",
    symmetric = TRUE,
    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
  ) +
  scale_y_sqrt() +
  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
  theme_minimal()
```

 5. Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD (final measure). Harmony pair with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$.

## results on cricket data

Gamma

```{r}
library(readr)
read_rds("data/smart_harmony_normal_cric.rds")
```
Normal 
```{r}
read_rds("data/smart_harmony_gamma_cric.rds")
```

>>>>

---
title: "Choosing harmonies based on average maximum pairwise distance of distributions"
output:
  html_document:
    theme: united
    highlight: tango
    toc: true
---

@@

  a) Normal: Pekalska and Duin [23] assumed that Lp-distances between
feature vectors are normally distributed, if the feature values are independent, identically distributed, and have limited variance. (“The Dissimilarity Representation for Pattern Recognition | Series in Machine Perception and Artificial Intelligence.” n.d. Accessed May 8, 2020. https://www.worldscientific.com/worldscibooks/10.1142/5965.)
We can not assume that feature values are independent.

  b) Gamma: (Ferencz, A., E. G. Learned-Miller, and J. Malik. 2005. “Building a Classification Cascade for Visual Identification from One Example.” In Tenth IEEE International Conference on Computer Vision (ICCV’05) Volume 1, 1:286–93 Vol. 1.) have considered the Gamma distribution to model the L2-distances.
This is intuitive since if the reponse variable follows Normal distribution, then $L_2$ norm distance measure which is similar to taking squares of differences of Normal variables would follow a Chi-Squared distribution, a special case of Gamma distribution.

@@

<!-- # Algorithm -->

<!--  1. Suppose $C_1$ and $C_2$ are two cyclic granularities such that $C_1$ maps index set to a set $\{A_1, A_2, A_3, \dots, A_I$\}, and $C_2$ maps index set to a set $\{B_1, B_2, B_3, \dots, B_J$\} and $v$ is the measured variable. Hence for each combination ($A_i$, $B_j$), we have s subset of the time series variable $v_{ij} \subseteq v$, $\forall i = {1, 2, \dots, I}$ and $\forall j = {1, 2, \dots, J}$ -->

<!--  2. There will be $J$ groups corresponding to each level $A_i$ of $C_1$. We compute the pairwise differences in distributions of these $J$ groups. Thus there will be  $k = \binom{J}{2}$ pairwise distances corresponding to each $A_i$. Let $D_1$,...,$D_k$ be these distances and let their maximum be $M_k$. -->

# Smart meter data

## Distances follow Weibull

```{r}
library(readr)
read_rds("data/smart_harmony_weibull.rds") %>% knitr::kable()
```

## Distances follow Gamma

 <!-- 3.1. Suppose the reponse variable follows Normal distribution. Then any distance measure which is similar to taking squares of differences of Normal variables would follow a Chi-Squared distribution, a special case of Gamma distribution. Hence if we proceed with the assumption that the $D_i$ distribution has Gamma distribution, then under EVT the maximum converges to Gumbel distribution and lim $P(M_k - b_k)/a_k)$ converges to a limiting Gumbel distribution where $b_k = (\frac{1}{\beta} (ln n + (\alpha - 1)ln ln n - ln \gamma(\alpha))$ and $a_k = \frac{1}{\beta}$. Here $\alpha$ and $\beta$ could be estimated using $(\mu/s)^2$ and $(s^2/\mu)$ respectively. -->

 <!-- 4.1. For each facet, we have a normalised variable using Step 3.1. We can take the maximum across all facets as the final mesasure. But this maximum is not normalized against the number of facet levels. Or we can take the mean (to be discussed). -->

```{r}
read_rds("data/smart_harmony_gamma.rds") %>% knitr::kable()
```

## Distances follow Normal

<!--  3.3. Let $D_{k} \sim N(\mu, \sigma^2)$. Then for $a_k = \mu − \sigma\Phi^{-1}(1/k)$ and $b_k = -\sigma/\Phi^{-1}(1/k)$, we have we have $P(\frac{M_k - a_k}{b_k} \leq z) \xrightarrow{} G(z)$, where $G$ is the standard Gumbel distribution. So normalization could be done  using $\frac{M_k - a_k}{b_k}$, where $\Phi^{-1}(1/k)$ could be approximated by $(1 - \frac{1}{k})$ sample quantile and $\mu$ and $\sigma$ by sample mean and standard deviation.  -->

<!--  4.3. We have been able to normalize for the number of levels of x-axis variable using Step 3.2. Now  -->
<!-- $\forall A_i, \quad i \in \{1, 2, \dots, I\}$, we have a normalised variable each of which follows a standard Gumbel distribution. Now we need to combine these normalised variables across different facets in a way so that the number of facet levels do not play a role. For this we can take the mean, median or any quantiles of these normalised variables. But the median (and any quantiles) of these Gumbel distributions grow at the rate of (some power of) $log n$ with $n \xrightarrow{} \infty$. Hence, median($N_i$)/log($I$) could serve as the final measure. -->

```{r}
library(readr)
read_rds("data/smart_harmony_normal.rds") %>% knitr::kable()
```

```{r, echo = FALSE, eval = FALSE}
smart_meter10 %>%
  filter(customer_id %in% c(10017936)) %>%
  prob_plot("hour_day",
    "wknd_wday",
    response = "general_supply_kwh",
    plot_type = "quantile",
    symmetric = TRUE,
    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
  ) +
  scale_y_sqrt() +
  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
  theme_minimal()
```

<!--  5. Steps 1 to 5 are repeated for every harmony pair and harmony pairs rearranged from highest to lowest MMPD (final measure). Harmony pair with higher MMPD would be more interesting as it would mean that the harmony pair exhibits more variations in distributions on an average between different levels of $C_2$. -->

# Cricket data

## Distances follow Weibull
```{r}
read_rds("data/smart_cric_weibul.rds")%>% knitr::kable()
```

## Distances follow Gamma

```{r}
library(readr)
read_rds("data/smart_cric_gamma.rds") %>% knitr::kable()
```

## Distances follow Normal 
```{r}
read_rds("data/smart_cric_normal.rds")%>% knitr::kable()
```

# Old benchmarks

@@

```{r, eval=FALSE, echo=FALSE}
read_rds("data/smart_harmony_ak.rds")
```

```{r, eval = FALSE}
smart_meter10 %>%
  filter(customer_id %in% c(10017936)) %>%
  prob_plot("day_week",
    "day_month",
    response = "general_supply_kwh",
    plot_type = "quantile",
    symmetric = TRUE,
    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
  ) +
  scale_y_sqrt() +
  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
  theme_minimal()
```

@@

```{r, eval = FALSE, echo=FALSE}
library(readr)
read_rds("data/smart_harmony_n.rds") %>% knitr::kable()
```

```{r, echo = FALSE, eval = FALSE}
smart_meter10 %>%
  filter(customer_id %in% c(10017936)) %>%
  prob_plot("week_month",
    "wknd_wday",
    response = "general_supply_kwh",
    plot_type = "quantile",
    symmetric = TRUE,
    quantile_prob = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99)
  ) +
  scale_y_sqrt() +
  #ggtitle(" (b) Area quantile plot faceted by weekend-weekday") +
  ylab("electricity demand [KWh]") + xlab("hours of the day") + ggtitle("") + ylab("")+ 
  theme_minimal()
```
